---
title: "R Plots"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
author: "Ethan Haley"
---

## Principles of Data Visualization and Introduction to ggplot2

**I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. Let's read this in:**

```{r, include=F}
library(dplyr)

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture1/Data/inc5000_data.csv", header= TRUE)
```

**And let's preview this data:**

```{r}
head(inc)
```
```{r}
summary(inc)
```

**Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:**

### How are the growth companies distributed, by Industry?

```{r}
# Insert your code here, create more chunks as necessary
inc %>%
  select(Industry) %>%
  table() %>%
  sort() %>%
  rev()
```

And how are they distributed by State?

```{r}
inc %>% 
  select(State) %>% 
  table() %>% 
  sort() %>% 
  rev()
```


Which Industry has the highest median growth rate?  

```{r}
inc %>%
  group_by(Industry) %>%
  summarise(MedianGrowth=median(Growth_Rate)) %>%
  arrange(desc(MedianGrowth))
```

Possible talking points:  

- The median growths are very similar per industry:  about 1% - 2%

- There are huge outliers in all 3 numericals (Growth Rate, Revenue, Employees)

- Companies per state would be more insightful if scaled by something like state population

- This is a list of some 5000 companies, so it might be more useful to weight big companies vs. small

## Question 1

**Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.**




```{r}
# Answer Question 1 here
library(ggplot2)
# Use a paper size as aspect ratio
PORTRAIT = 11 / 8.5
# Possibly limit overcrowding of chart
NSTATES = 52   

states = inc %>% 
  select(State) %>% 
  table() %>%
  sort() %>% 
  tail(NSTATES) %>%
  as.data.frame()

states %>%
  ggplot(aes(x = ., y = Freq, width=.5)) +
  geom_col() +
  coord_flip() +
  ggtitle('Fast-Growth Company Locations') +
  theme_minimal() +
  theme(aspect.ratio = PORTRAIT) +
  ylab('Number of Companies') +
  theme(axis.text.y = element_text(size = 6))
```

Depending on the audience and topic, it could be more meaningful to scale these numbers by total number of companies in the state.


## Question 2

**Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.**

```{r}
# inspect numbers
inc %>%
  filter(State=='NY') %>%
  filter(complete.cases(.)) %>%
  group_by(Industry) %>%
  summarise(MedianEmpl=median(Employees), Companies=n()) %>%
  arrange(desc(MedianEmpl))
```

```{r}
# Answer Question 2 here

inc %>%
  filter(State=='NY') %>%
  filter(complete.cases(.)) %>%
  ggplot(aes(reorder(Industry, Employees, FUN=median), Employees)) +
  geom_boxplot(varwidth = T, outlier.size = .7) + 
  coord_flip() +
  theme_minimal() +
  xlab('') +
  scale_y_log10() +
  labs(title = 'Employees per NY Growth Company, by Industry',
       caption='Log-scaled, and with bar thickness proportional to number of companies in that industry') 

```

Notes: 

- The box thickness is actually proportional to the square root of the number of companies in that industry.  

## Question 3

**Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.**

```{r}
# Answer Question 3 here

inc %>%
  filter(complete.cases(.)) %>%
  mutate(RpE = Revenue / Employees) %>%
  ggplot(aes(reorder(Industry, RpE, FUN=median), RpE)) +
  geom_boxplot(varwidth = T, outlier.size = .7) + 
  coord_flip() +
  theme_minimal() +
  xlab('') +
  scale_y_log10() +
  labs(title = 'Revenue per Employee at each Company, by Industry',
       caption='Log-scaled, and with bar thickness proportional to number of companies in that industry') 

```

Notes:  

- I interpret this question to include all states, not just NY.  Comparing box thicknesses in this plot to the previous one, you can see the outsized representation of Advertising & Marketing companies in NY, compared to relatively lower NY representation by IT Services, for example.  At least in terms of company counts.  




